---
title: "COVID-19 Data from NYT for Counties of Interest"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Updated: `r date()`

## Get the Data

The NYT is super generous with their data. Go get it.

```{r data}
library(tidyverse)
library(magrittr)
curl::curl_download(
  url      = 'https://github.com/nytimes/covid-19-data/raw/master/us-counties.csv',
  destfile = './county-cases.csv'
)
df <- read_csv('county-cases.csv') 
```

Are you curious what it looks like? It looks like this.

```{r showdata, results = 'asis'}
df %>%
  sample_n(5) %>%
  pander::pandoc.table(caption = 'NYT COVID Tracking Data')
```

## Tidy the data

Tidy the data by filtering on the counties of interest, and gathering values of interest into a single column.

```{r tidy}
county_list = c(
  'Alameda, California', 
  #'Contra Costa, California', 
  #'Bernalillo, New Mexico', 
  #'Mecklenburg, North Carolina',
  'Mercer, New Jersey'
)
df %<>%
  # filter on the county_list
  mutate(county = paste(county, state, sep = ', ')) %>%
  # fips is NA when the data is at the state level
  filter(!is.na(fips)) %>%
  left_join(select(usmap::countypop, fips, population = pop_2015), 
            by = 'fips') %>%
  filter(county %in% county_list) %>% 
  # we don't need everything, keep only the region, day, and counts
  select(county, date, cases, deaths, population) %>%
  # tidy the data by putting all observations in rows
  gather(key = type, value = number, -date, -county, -population) %>%
  # group by county and type (deaths or cases) to get a delta column for new cases in a day
  # cases and deaths should be monotonically increasing, but the data is imperfect. 
  # correct using the following lines, assuming that sometimes the count is prematurely high
  group_by(county, type) %>%
  arrange(date %>% desc) %>%
  mutate(number = cummin(number)) %>%
  arrange(date) %>%
  mutate(new = number - lag(number, default = 0)) %>%
  ungroup()
```

Curious what it looks like now? It looks like this.

```{r tidy-show, results='asis'}
df %>%
  arrange(date) %>%
  sample_n(5) %>%
  pander::pandoc.table(caption = 'Tracking Data Ready for Plotting')
```

## Plot the data

This is what the straight-up cumulative data looks like.

```{r plot-cumulative, fig.height=3, fig.width=6}
# time for plotting
library(ggplot2)
ggplot(df, aes(x = date, y = number/population*1e5, fill = type)) +
  geom_area(alpha    = 0.5,
            position = 'identity',
            show.legend = FALSE) +
  scale_y_log10(name   = 'Per 100,000 People',
                labels = scales::comma) +
  scale_x_date(name        = '',
               date_breaks = '1 week',
               labels      = scales::date_format('%m/%d')) +
  geom_hline(yintercept = 10^(-2:3), size = 0.5, color = 'white') +
  facet_grid(type~county, scales = 'free_y') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 50, hjust = 1))
```

This looks at the daily increase.

```{r plot-daily, fig.height=3, fig.width=6}
ggplot(df, aes(x = date, y = new/population*1e5, color = type)) +
  geom_hline(yintercept = 10^(-2:3), size = 0.1) +
  geom_step(show.legend = FALSE) +
  scale_x_date(name        = '',
               date_breaks = '1 week',
               labels      = scales::date_format('%m/%d')) +
  scale_y_log10(name   = 'Daily Cases Per 100,000 People', 
                breaks = 10^(-2:3),
                limits = c(0.01,1000),
                labels = scales::comma) +
  facet_grid(type~county, scales = 'free_y') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 50, hjust = 1))
```

Because of the noise in the daily data, maybe a weekly increase is a better idea.
```{r plot-weekly, fig.height=3, fig.width=6}
df %>% 
  dplyr::group_by(lubridate::week(date), type, county, population) %>% 
  dplyr::summarize(new = sum(new), date = min(date)) %>%
  ggplot(aes(x = date, y = new/population*1e5, color = type)) +
  geom_hline(yintercept = 10^(-2:3), size = 0.1) +
  geom_step(show.legend = FALSE) +
  scale_x_date(name        = '',
               date_breaks = '1 week',
               labels      = scales::date_format('%m/%d')) +
  scale_y_log10(name   = 'Weekly Cases Per 100,000 People', 
                breaks = 10^(-2:3),
                limits = c(0.01,1000),
                labels = scales::comma) +
  facet_grid(type~county, scales = 'free_y') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 50, hjust = 1))
```